Analysis of Red Wine Chemical Attributes by Javier Ortiz
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
library(ggplot2)
library(knitr)
library(dplyr)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
reds <- read.csv('wineQualityReds.csv')
head(reds)
nrow(reds)
```

The Red Wine Quality data set contains information about over 1,500 red wines
with 11 variables that describe the chemical characteristics of wine. The 
data set provides a rating of the quality of each wine, with 0 being the 
lowest quality and 10 being the greatest quality.

Some questions that would be relevant to this dataset are what chemical
characteristics do quality wines share, do wines
with higher alcohol content receive better ratings, do sweeter wines receive 
better ratings, what level of acidity is associated with higher quality wines.

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
ggplot(aes(x = fixed.acidity), data = reds) + 
  geom_histogram(binwidth = .1) +
  scale_x_continuous(breaks = seq(4, 17, 1))
```

The histogram for fixed acidity is right skewed. Most of the observations fall
in the 7-8 range.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = fixed.acidity), data = reds) + 
  geom_histogram(binwidth = .1) +
  scale_x_continuous(breaks = seq(4, 17, 1)) +
  facet_wrap(~quality)
```

I was hoping to find a different distribution of fixed acidity for higher
quality wines, but the distributions for all the wine qualities seem similar 
with most observations falling in the 6 to 11 range.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = volatile.acidity), data = reds) + 
  geom_histogram(binwidth = .01) +
  scale_x_continuous(breaks = seq(0, 1.6, .1))
```

The distribution for volatile.acidity is more disperse, but still, most
of the observations fall in the lower half of the values, between .3 and .7 with
some outliers having values around 1.6.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = volatile.acidity), data = reds) + 
  geom_histogram(binwidth = .01) +
  scale_x_continuous(breaks = seq(0, 1.6, .1)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~quality)
```

Volatile acidity is the amount of acetic acid in wine which according to the
text file that comes with the data set, at too high levels might give an
unpleasant taste to the wine. We can notice that higher quality wines do have
a slighltly smaller value for volatile acidity in comparison to the lesser
quality wines.

```{r echo=FALSE, message=FALSE, warning=FALSE}
by(reds$volatile.acidity, reds$quality, summary)
```

There seems to be a trend for volatile acidity, that decreases as the wine
quality improves. Wine qualities of 7 and 8 have the lowest medians for this
value with .3700, while lower quality wines have a median of .8450.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = citric.acid), data = reds) + 
  geom_histogram(binwidth = .01) +
  scale_x_continuous(breaks = seq(0, 1, .10))
```

Even though most of the observations fall in the lower half of the distribution,
there is one clear peak around the value of 0 and another less pronounced 
peak at around the value of 0.5.

```{r echo=FALSE, message=FALSE, warning=FALSE}
by(reds$citric.acid, reds$quality, summary)
```

There is a clear trend with the means and medians for citric acid, which
increases as the wine quality increases. The lowest quality wines have a mean
of .1710 and the highest quality wines a mean of .3911.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = residual.sugar), data = reds) + 
  geom_histogram(binwidth = .1) +
  scale_x_continuous(breaks = seq(0, 16, 1))
```

The histogram for residual sugar is right skewed. Most of the observations have
a value of 2.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = residual.sugar), data = reds) + 
  geom_histogram(binwidth = .1) +
  scale_x_continuous(breaks = seq(0, 16, 1)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~quality)
```

It doesn't seem that residual sugar is a distinctive characteristic of higher
quality wines.

```{r echo=FALSE, message=FALSE, warning=FALSE}
by(reds$residual.sugar, reds$quality, summary)
```

All qualities of wine have a very close mean for residual sugar.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = chlorides), data = reds) + 
  geom_histogram(binwidth = .01) +
  scale_x_continuous(breaks = seq(0, .6, .05))
```

The distribution for chlorides is heavily skewed to the right, with the vast
majority of observations falling in the .05 and .1 range.

```{r echo=FALSE, message=FALSE, warning=FALSE}
by(reds$chlorides, reds$quality, summary)
```

It turns out that high quality wines have the lowest mean for the value of 
chlorides with .06844 while low quality wines have a mean of .1225.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = chlorides), data = reds) + 
  geom_histogram(binwidth = .01) +
  scale_x_continuous(breaks = seq(0, .6, .05)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~quality)
```

The distribution for chlorides is very similar for all the different
wine qualities.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = free.sulfur.dioxide), data = reds) + 
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 74, 2)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The distribution for free sulfur dioxide is right skewed, with most of 
the observations falling between the 4 and 18 range. There are some outliers far
above the value of 58.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = total.sulfur.dioxide), data = reds) + 
  geom_histogram(binwidth = 5) +
  scale_x_continuous(breaks = seq(0, 300, 20))
```

The distribution for total sulfur dioxide is right skewed with most of the
observations falling in the 15 - 25 range. There are some outliers around the
280 value.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = total.sulfur.dioxide), data = reds) + 
  geom_histogram(binwidth = 5) +
  scale_x_continuous(breaks = seq(0, 300, 20)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~quality)
```

All the rankings for quality have a small number for total sulfur dioxide,
however there are some outliers with high values for wine quality number 7.

```{r echo=FALSE, message=FALSE, warning=FALSE}
by(reds$total.sulfur.dioxide, reds$quality, summary)
```

I thought that a low level of total sulfur dioxide might be an attribute of 
high quality wines, but bad quality wines actually have a lower median and mean
for this value.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density), data = reds) + 
  geom_histogram(binwidth = .0002) +
  scale_x_continuous(breaks = seq(.990, 1.005, .002))
```

The observations for density are normally distributed with most values being 
between .996 and .998.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = pH), data = reds) + 
  geom_histogram(binwidth = .02) +
  scale_x_continuous(breaks = seq(2.7, 4, .1))
```

The pH variable follows a normal distribution with most of the observations
falling between the 3.2 and 3.4 values.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = sulphates), data = reds) + 
  geom_histogram(binwidth = .02) +
  scale_x_continuous(breaks = seq(.3, 2, .1))
```

The distribution for sulphates is right skewed, with most of the observations
falling in the .5 - .7 range, with some outliers near the value of 2.

```{r echo=FALSE, message=FALSE, warning=FALSE}
by(reds$sulphates, reds$quality, summary)
```

When we calculate the mean sulphates for every quality ranking, we can notice
that the mean sulphate level increases as the quality of wine improves.
Low quality wines have a sulphates mean of .5700, while high quality wines
have a mean of .7678.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = alcohol), data = reds) + 
  geom_histogram(binwidth = .10) +
  scale_x_continuous(breaks = seq(8, 16, 1))
```

The alcohol content varies between 8 and 15, with most observations falling
between 9.20 and 10.5.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = alcohol), data = reds) + 
  geom_histogram(binwidth = .10) +
  scale_x_continuous(breaks = seq(8, 16, 1)) +
  facet_wrap(~quality)
```

When we look at the alcohol levels by each quality ranking we can notice that
most wines have a quality of 5. The lowest quality wines also have the lowest
alcohol content, under 11, while the highest quality wines, have an alcohol
content above 11. The middle quality wines have a more disperse alcohol 
content between 9 and 12.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality), data = reds) + 
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(1, 10, 1))
```

The wine quality follows a normal distribution with most wines being a 5 or a 6.
It seems that not a single wine got a score higher than 8 or lower than 3. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
table(reds$quality)
```

Most wines have a quality ranking of 5 or 6, with only 10 wines in the dataset
being really bad (a ranking of 3) and 18 being really good (a ranking of 8).

```{r echo=FALSE, message=FALSE, warning=FALSE}
by(reds$alcohol, reds$quality, summary)
```

There is definetely a trend in the means and medians of alcohol content that
increases as the quality improves. Wines with a quality of 3 have an
alcohol mean of 9.925 while wines with a quality of 8 have an alcohol content
mean of 12.09.

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(reds$alcohol)

# This code converts the alcohol variable into a factor.
reds$alcohol_ranking <- ifelse(reds$alcohol <= 9.50, 'low',
                               ifelse(reds$alcohol <= 10.20,
                                      'medium-low',
                                      ifelse(reds$alcohol <= 11.10,
                                             'medium-high', 'high')))
reds$alcohol_ranking <- factor(reds$alcohol_ranking)
head(reds)
```

Here I created a new variable that classifies each wine as
low, medium-low, medium-high or high according to the distribution
of quantiles in the alcohol variable.

```{r}
ggplot(aes(x = alcohol_ranking), data = reds) + 
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 500, 50))

table(reds$alcohol_ranking)
```

This graph shows the count for each alcohol ranking. Most wines (436) have a 
ranking of low alcohol, which is below 9.50, followed by the medium-low
category, with 406 observations. The high and medium-high rankings followed
behind with 380 and 377 observations each.

# Univariate Analysis

### What is the structure of your dataset?
The data set is made of 1599 observations of red wines with 11 variables
that describe their chemical properties.

### What is/are the main feature(s) of interest in your dataset?
The main feature of interest in the dataset is the quality ranking for each
wine. It might be really interesting to find out what factors affect
the quality of wine the most.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
I think alcohol content, volatile acidity, citric acid, chlorides and 
sulphates might be the chemical properties that could have the most weight
when determining the quality of wine.

### Did you create any new variables from existing variables in the dataset?
I created the variable alcohol_ranking, using the quantiles of the
alcohol sample, I classified each wine either as low, medium low,
medium high, or high. Finally, I converted that variable into a
factor because that will be really useful when performing
multiple variable analysis.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
Most distributions were right skewed. Just density, pH and quality followed a
normal distribution. I didn't feel the need to change the form of the data,
the only modification done was calculating mean, median and quantiles of 
certain features in order to understand the distributions better.

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
reds$quality_factor <- as.factor(reds$quality)

ggplot(aes(x = quality_factor, y = alcohol), data = reds) + 
  geom_boxplot(alpha = 1/4)
```

Low quality wines have a small alcohol content mean, with quality #3
having a value around 10, quality 4 moving upwards a little bit, and
quality 5 going back down under quality #3, but after that, there
is a consistent linear increment in the alcohol content mean as the
quality improves, reaching the peak of the distribution at the
highest quality.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality_factor, y = volatile.acidity), data = reds) +
  geom_boxplot(alpha = 1/4)
```

This time we have an inverted trend where volatile acidity decreases as
the quality improves. We can notice that high quality wines have a much 
smaller distribution, with most observations being around the value of 
.4 for volatile acidity, which is not the case for other wine qualities 
that have a more spread out distribution.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality_factor, y = citric.acid), data = reds) +
  geom_boxplot(alpha = 1/4)
```

For citric acid, there is a clear trend that shows how citric acid increases as
the quality of wine improves.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality_factor, y = residual.sugar), data = reds) +
  geom_boxplot(alpha = 1/4)
```

It doesn't seem that residual sugar is likely to have an impact on
quality.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality_factor, y = chlorides), data = reds) +
  geom_boxplot(alpha = 1/4)
```

It does seem that higher quality wines usually have a smaller
value for chlorides in comparison to lower quality wines, but the differnce
is small.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality_factor, y = sulphates), data = reds) +
  geom_boxplot(alpha = 1/4)
```

There seems to be a trend where as the wine quality improves,
the wine has a higher level of sulphates.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# This creates another dataset that groups alcohol mean and median according to
# quality.
reds.alcohol_by_quality <- reds %>%
  group_by(quality) %>%
  summarise(alcohol_mean = mean(alcohol),
            alcohol_median = median(alcohol),
            n = n()) %>%
  arrange(quality)

head(reds.alcohol_by_quality)
```

Here we can see the alcohol mean, alcohol median and number of 
observations for each wine quality. Higher quality wines have
a higher alcohol mean and alcohol median. Most wines have a 
quality of 6 or 7.

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(reds$quality, reds$alcohol, method = 'pearson')
```

Running a correlation test we get a result of .476 which is a moderate 
but meaningful correlation.

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(reds$quality, reds$volatile.acidity, method = 'pearson')
```

After a correlation test between quality and volatile acidity,
we get the result -0.391 which is a small but meaningful correlation.

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(reds$quality, reds$citric.acid, method = 'pearson')
```

After a correlation test between quality and citric acid,
we get a result of .226 which indicates that there is not
a meaningful relationship between the two variables.

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(reds$quality, reds$chlorides, method = 'pearson')
```

After a correlation test between quality and chlorides,
we get a result of -0.129, which indicates that there
is no meaningful relationship between these two variables.

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(reds$quality, reds$sulphates, method = 'pearson')
```

After a correlation test between quality and sulphates,
we get a result of .251, which indicates that there is
no meaningful relationship between these two variables.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = fixed.acidity, y = volatile.acidity), data = reds) +
  geom_point(alpha = 1/4)
```

The relationship between fixed acidity and volatile acidity is not very clear.
It seems that most observations have a fixed and volatile acidity in the 
lower half of the range.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = fixed.acidity, y = residual.sugar), data = reds) +
  geom_point(alpha = 1/4)
```

Most wines have a residual sugar level that is less than 4, even across the 
different levels of fixed acidity. I thought that maybe acidity could be
used to balance the sweetness of a wine, so I was hoping to find that wines
with a higher level of fixed acidity would be less sweet. This chart doesn't
tell that narrative.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density, y = chlorides), data = reds) +
  geom_point(alpha = 1/4)
```

There seems to be a really small trend when chlorides increase so does the 
density of wine.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = residual.sugar, y = alcohol), data = reds) +
  geom_point(alpha = 1/4)
```

I have heard before that really sweet wines usually have more alcohol.
This graph discredits that belief. Most wines, no matter how much
alcohol they have, are in the low end of the spectrum regarding
residual sugar. Among those outliers with a lot of residual sugar,
most of them fall in the low alcohol content category.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#This converts quality into a factor
ggplot(aes(x = quality_factor, y = alcohol), data = reds) +
  geom_boxplot(alpha = 1/4)
```

This graph clearly shows how higher quality wines have a 
higher alcohol content, not only a higher mean or median, but
the entire distribution is above the values that lower quality
wines have.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
Some of the relationships I observed were about quality against different
features: alcohol, volatile acidity, chlorides, sulphates and citric acid.
Even though most of the scatterplots showed a trend that links these various
features with quality, the correlations were non-significant and the trends
too weak, with the exception of alcohol and quality.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
Yes, I had the idea that really sweet wines had a higher alcohol content. 
The scatterplot not only debunked this myth, since most wines are not that
sweet regardless of their alcohol content, but out of those wines that are
sweet, they usually have a low alcohol content.

### What was the strongest relationship you found?
The strongest relationship I found is between alcohol and quality. The 
correlation test came out with a result of 0.476 which is a moderate but
significant correlation, and when comparing the alcohol content means
across quality, the high quality wines have a signficant higher alcohol
content than low quality wines.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(aes(x = volatile.acidity, y = alcohol), data = reds) +
  geom_point(alpha = 1/4) +
  facet_wrap(~quality_factor)
```

Here we have a graph for each quality ranking that describes the  relationship 
between alcohol and volatile.acidity. Low quality wines have a low alcohol 
content and a relatively high volatile acidity. On the other hand, 
high quality wines have a higher alcohol content and a lower volatile acidity.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = citric.acid, y = alcohol),
       data = reds) +
  geom_point(aes(color = quality_factor)) +
  scale_colour_brewer(palette = "RdYlGn") +
  theme_dark()
```

This graph shows the relation of alcohol and citric acid
for each quality level. High quality wines have a higher
alcohol content with medium levels of citric acid.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = alcohol, y = pH), data = reds) +
  geom_point(aes(color = quality_factor)) +
  scale_colour_brewer(palette = "RdYlGn") +
  theme_dark()
```

Nothing particular stands out.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = residual.sugar, y = alcohol), data = reds) +
  geom_point(aes(color = quality_factor)) +
  scale_colour_brewer(palette = "RdYlGn") +
  theme_dark()
```

This graph confirms that the sweetest wines have the lowest alcohol
content and also that high quality wines tend to have low levels
for residual sugar.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density, y = chlorides), data = reds) +
  geom_point(aes(color = alcohol_ranking)) +
  facet_wrap(~quality_factor) +
  scale_colour_brewer(palette = "RdYlGn") +
  theme_dark()
```

Wines with a low alcohol ranking (green dot) are the ones
that have higher values for chlorides.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = volatile.acidity, y = citric.acid), data = reds) +
  geom_point(aes(color = quality_factor)) +
  scale_colour_brewer(palette = "RdYlGn") +
  theme_dark()
```

There is a clear trend that establishes a relationship between citric
acid and volatile acidity. When levels of citric acid are high, volatile
acidity tends to be low, this is specially true in high quality wines.
The opposite is also true, when volatile acidity is high, citric acid
tends to be low, this happens more often with lower quality wines.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = residual.sugar, y = volatile.acidity, color = alcohol_ranking), 
       data = reds) +
  geom_point() +
  facet_wrap(~quality_factor) +
  scale_colour_brewer(palette = "RdYlGn") +
  theme_dark()
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
For this section, I already had a clear idea that alcohol content was the 
most important feature in determining the quality of wine. So I focused
my attention on describing the relationship of alcohol with other
features of the dataset. High quality wines (which have high alcohol
content), have low levels for volatile acidity and medium levels
for citric acid. 
Another interesting relationship is that of high quality wines and 
residual sugar, that shows that high quality wines have a low content
of residual sugar, thus describing a high quality wine that is 
high in alcohol content and dry regarding its residual sugar. 

### Were there any interesting or surprising interactions between features?
Yes, the most interesting being the relationship between chlorides, 
alcohol content and quality. It turns out that levels of chlorides remain
steady across the dataset with the exception of low alcohol wines, that
have a clear spike in their levels of chlorides, which might suggest
a relation between chlorides and alcohol content. Another relation that
derives from the previous one is that this low alcohol - high chlorides
wines appear exclusevily in the levels of quality 5 and 6.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(aes(x = quality_factor, y = alcohol), data = reds) +
  geom_boxplot(outlier.alpha = 1/4) +
  scale_y_continuous(breaks = seq(8, 15, 1)) +
  ggtitle('Alcohol vs Quality') +
  labs(x = 'Quality (score between 0 and 10)', y = 'Alcohol (% by volume)')
```

### Description One
This graph shows the strongest relationship I found in the dataset: Alcohol and
Quality. If we follow the means of each quality distribution, we can see a clear
trend that starts at quality level 5, where the alcohol content increases as 
the quality improves. As a result, wines with a quality of 6 have a higher 
alcohol content than wines with a quality of 5, wines with a quality of 7 have 
even higher alcohol content, and wines with a quality of 8 have the highest 
alcohol content of the sample. It is also clear that for the highest quality 
wines with a quality ranking #8, the bulk of the distribution (interquantile 
range) has a higher alcohol content than that of the entire distribution of 
wines with quality ranking #3.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
ggplot(aes(x = residual.sugar, y = alcohol, color = quality_factor),
       data = reds) +
  geom_point() +
  scale_color_brewer(type = 'qual',
                     guide = guide_legend(title = 
                                            'Quality', 
                                          reverse = TRUE)) +
  scale_y_continuous(breaks = seq(8, 16, 1)) +
  scale_x_continuous(breaks = seq(0, 16, 1)) +
  ggtitle('Alcohol vs Residual Sugar of Quality') +
  labs(x = 'Residual Sugar (g / dm^3)', y = 'Alcohol (% by volume)') +
  theme_dark()
```

### Description Two
I like this plot because it weakens the myth that sweet wines have a high
alcohol content. Most wines, regardless of their quality, have a residual
sugar level of 4g/l, that categorizes them as dry. This is specially true
with high quality wines with a ranking of 8, since most of the distribution
is below this parameter. It is also clear that quality ranking #7  and #5
have the highest number of observations in the medium-dry range which
covers from 4g/l up to 12g/l. We can only observe a couple of wines with a 
quality of 8 in this category. Finally, in the sweetest wines in this plot,
above 12g/l which is considered a medium sweetness, we don't have any
observations of quality #8 wines.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x = citric.acid, y = volatile.acidity, color = alcohol_ranking),
       data = reds) +
  geom_point() +
  scale_y_continuous(breaks = seq(0, 1.6, .2)) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 
                                            'Alcohol Ranking')) +
  ggtitle('Citric Acid vs Volatile Acidity of Quality and Alcohol') +
  labs(x = 'Citric Acid (g / dm^3)', 
       y = 'Volatile Acidity (acetic acid - g / dm^3)') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~quality_factor)
```

### Description Three
This graph shows the relationship between volatile acidity and citric acid
by comparing the different alcohol rankings across the different quality
rankings. For low quality wines such as #3, is has become clear that they 
have a lower alcohol content since there are no observations of wines with
a high alcohol content in that quality ranking. Furthermore, low quality wines
seem to have a higher value for volatile acidity and lower values for citric
acid. When comparing quality rankings #3 against #8, this becomes obvious:
While quality ranking #3 has most of their observations with a value above
.8 for volatile acidity, quality ranking #8 has most of its observations
with a value below .6 for volatile acidity; while quality ranking #3 has
most of its observations with a value below .25 for citric acid, quality
ranking #8 has most of its observations with a value above .25 for citric
acid. All quality rankings follow the trend that when volatile acidity is 
high, citric acid is low; and when volatile acidity is low, citric acid
is high.

------

# Reflection

Here I list some of the key insights found through this exploration.

The sample follows a normal distribution for quality, with most wines having
a quality of 5 or 6. One surprising of the dataset is that there wasn't
a single observation below 3 or above a quality of 8.

Alcohol content proved to be one of the most important features when
determining if a wine is high quality. According to the sample, most high
quality wine have a high alcohol content while most of the low quality wines
have also a low alcohol content.

One interesting relationship is that between alcohol content and residual
sugar. Some people think that the sweeter the wine the more alcohol it has.
It turns out this is not true, but opposite is. Most of the wines with high
quality have lower levels of residual sugar, and from those wines that have 
a high level of residual sugar, they are usually pretty low in alcohol.

It also was really interesting to observe how the relation between different
variables changed in each quality category. For example, volatile acidity and
citric acid experience a constant trend throughout the data set where if 
volatile acidity increases, citric acid decreases and also the opposite. This
became more obvious with high quality wines that have a much lower level of
volatile acidity, thus, higher levels of citric acid.

One of the main struggles is quantity of variables that affect the quality of
wine. Since there are so many, any subtle difference could change the quality
of the wine dramatically, so it was really difficult to establish trends
among the entire sample.

In the future it could be really interesting to add the type of red wine
that is being graded, since that could change the rating dramatically since
one is looking for different characteristics when tasting a cabernet sauvignon 
to a zinfandel for example. Having some wines with a quality of 9 or 10
could also be really interesting and perhaps help establish stronger trends
in the data.
